---
title: "simpleRegress"
author: "Citlali Gonzalez"
date: "7/9/24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
For model lenght~neck, adjusted r^2 = 0.75

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## length based on age

```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```

adjusted $r^2$ for the age model is 0.51, this is less than the     
adjusted $r^2$ for neck, so this model is not as good as the neck model

## length based on chest

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```
adjusted $r^2$ for the chest model is 0.79, this has to do with the      
adjusted $r^2$ for chest therefore the model is as good as the chest model. 

## length based on headlength 

```{r}
headlengthModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlengthModel)
summary.lm(headlengthModel)
```

For model length~headlength, adjusted $r^2$ = 0.84

adjust $r^2$ for the head length model is 0.84, this is greater than.     
adjusted $r^2$ for neck, this model is better than the neck model.

## length based on headwidth 

```{r}
headwidthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwidthModel)
summary.lm(headwidthModel)
```

adjusted $r^2$ = 0.53,      
adjusted $r^2$ for Length~Headlength is better than length~headwidth 

## length based on sex

```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
abline(sexModel)
summary.lm(sexModel)
```

adjusted $r^2$ = -0.0084,      
adjusted $r^2$ length~headlength is .84 this shows that it is better than length~sex

## length based on weight 

```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```

adjusted $r^2$ for weight is 0.75, this is less than adjusted $r^2$ for headwidth therefore this model is not a good model for headwidth.     
adjusted $r^2$ for headlength is the highest model

### Conclusion

I conclude that models based on headlength and chest measurements tend to show the highest adjusted $r^2$ which gives the best predictions. These variables were the highest compared to others such as headwidth, age, sex, and weight. 


